Synthesizing Symmetric Lenses

### 2018-07-11 papers/SymmetricLenses/main.tex
minor edits to section 6

% begin evaluation
\section{Evaluation}
\label{sec:evaluation}
We have implemented our symmetric DNF lenses and the synthesis of them as an
extension to Boomerang. We added symmetric lens combinators to Boomerang, and
implemented all Boomerang's asymmetric lens combinators using a combination of
the symmetric lens combinators, as well as other symmetric lens combinators for
the asymmetric theory extensions present in Boomerang, like matching lenses and
quotient lenses. We also integrated our synthesis engine into Boomerang. It allows users to write synthesis tasks
alongside lens combinators, use those synthesis results in the manual writing of
new lenses, and use previously defined lenses in the synthesis of new lenses.

@@

Recall that compositional synthesis exploits user annotations to break a benchmark into a series of smaller synthesis steps. In our experience, compositional synthesis (SS mode) allows our system to scale to arbitrarily large and complex formats. However, SSNC mode, which uses no such annotations and synthesizes a complete lens all at once provides a useful experimental stress test for the system. Recall from Section

\begin{figure}
  \includegraphics{generated-graphs/times}
  \caption{Number of benchmarks that can be solved by a given algorithm in a
    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
    \BSOpt{} is the full bijective lens synthesis algorithm. \SSNCOpt{} is the
    symmetric synthesis algorithm without using a library of existing lenses.
    \BSNCOpt{} is the bijective synthesis algorithm without using a library of
    existing lenses. Our symmetric synthesis algorithm is able to complete all
    benchmarks in our benchmark suite in under 30 seconds, and is able to
    complete a majority of the benchmarks without using compositional
    synthesis.}
  \label{fig:times}
\end{figure}

For each benchmark in our benchmark suite, we ran every mode on it with a
timeout of 30 seconds, and average the result over 5 runs.
Figure~\ref{fig:times} summarizes the results of these tests. We find that our
algorithm is able to synthesize all of the benchmarks in our benchmark suite in
under 30 seconds. Without compositional synthesis, the synthesis algorithm is
able to solve 26 problem instances. Recall from Section~\ref{subsec:optimizations} that users may tune the termination metric. Success on this set of benchmarks requires we tune that termination codition (e.g. for some benchmarks, allowing the system flexibility to spend more time considering more of the search space). In Section~\ref{subsec:effect-of-heuristics}, we will consider the effect of fixing the termination metric.

@@

To compare to the existing bijective synthesis algorithm, we run our symmetric
synthesis algorithm on the original Optician benchmark suite. \afm{8} of these
benchmarks had to be slightly altered to work with our new system. We altered
these benchmarks by either providing additional examples termination parameters.
Without these alterations, the symmetric synthesis synthesized a lens that fit
the specification but not the intended lens.

@@

For each benchmark in our benchmark suite, we ran every mode on it with a
timeout of 30 seconds, and average the result over 5 runs.
Figure~\ref{fig:times_bijective} summarizes the results of these tests. On
average, \SSOpt{} took 2.1 times (\afm{3.5} seconds) longer to complete than
\BSOpt{}. The slowest completed benchmark for both synthesis algorithms is
\texttt{xml\_to\_augeas.boom}, a benchmark that converts arbitrary XML up to
depth 3 into a serialized version of the structured dictionary representation
used in Augeas. This benchmark takes 18.9 seconds for the symmetric synthesis
algorithm to complete, and 9.3 seconds the bijective synthesis algorithm to
complete.

@@

\subsection{The Effects of Our Heuristics}
\label{subsec:effect-of-heuristics}
We evaluate the usefulness of (1) our information theoretic metric, and (2) our
termination heuristic.  To this end, we run our program in 3 different modes:\\
\begin{tabulary}{\linewidth}{rL}
  \AnyOpt{}: & We synthesize lenses without use of the information
               theoretic preference metric.  Once \GreedySynth finds a lens, the
  synthesis procedure returns that lens.\\
  \FLOpt{}: &  The first time \GreedySynth finds a lens, the synthesis procedure
              returns that lens. However, \GreedySynth uses the information
              theoretic preference metric to choose a lens.\\
  \CCOpt{}: & We replace our information theoretic cost metric with one where
              the cost of the lens is the number of disonnects plus the number
              of merges.\\
  \ZTOpt{}: & \ZTOpt{} runs the algorithm with the termination parameter
              universally set to 0 (default). \\
  \TFTOpt{}: & \TFTOpt{} runs the algorithm with the
               termination parameter universally set to 25. \\
  \NTFTOpt{}: & \NTFTOpt{} runs the algorithm with the
                termination parameter universally set to -25. \\
\end{tabulary}

@@

We experimented with the \CCOpt{} mode to determine whether the complexity of the information theoretic measure was really needed. Related work on string transformations has often used simpler measures such as ``avoid constants'' which align with, but are simpler than our measures. The \CCOpt{} mode is an example of such a simple measure --- it operates by counting disconnects, which put a complete stop to information transfer, like constants, and merges, which eliminate the information in a union. Figure~\ref{fig:metric} summarizes the result of this experiment. We find that
our metrics are very important for finding the correct lens. Our
information-theoretic metric is critical for finding the correct lens. Indeed,
only \afm{10} of our benchmarks succeeded when running in \CCOpt{} mode.

>>>>

% begin evaluation
\section{Evaluation}
\label{sec:evaluation}
We have implemented our symmetric DNF lenses and the synthesis of them as an
extension to Boomerang. We added symmetric lens combinators to Boomerang, and
implemented all of Boomerang's asymmetric lens combinators using a combination of
the symmetric lens combinators, as well as other symmetric lens combinators for
the asymmetric theory extensions present in Boomerang, like matching lenses and
quotient lenses. We also integrated our synthesis engine into Boomerang. It allows users to write synthesis tasks
alongside lens combinators, use those synthesis results in the manual writing of
new lenses, and use previously defined lenses in the synthesis of new lenses.

@@

Recall that compositional synthesis exploits user annotations to break a
benchmark into a series of smaller synthesis steps. In our experience,
compositional synthesis (SS mode) allows our system to scale to arbitrarily
large and complex formats. However, SSNC mode, which uses no such annotations
and synthesizes a complete lens all at once provides a useful experimental
stress test for the system.

\begin{figure}
  \includegraphics{generated-graphs/times}
  \caption{Number of benchmarks that can be solved by a given algorithm in a
    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
    \saz{BS isn't in the figure}\BSOpt{} is the full bijective lens synthesis algorithm. \SSNCOpt{} is the
    symmetric synthesis algorithm without using a library of existing lenses.
    \saz{BSNC isn't in the figure}\BSNCOpt{} is the bijective synthesis algorithm without using a library of
    existing lenses. Our symmetric synthesis algorithm is able to complete all
    benchmarks in our benchmark suite in under 30 seconds, and is able to
    complete a majority of the benchmarks without using compositional
    synthesis.}
  \label{fig:times}
\end{figure}

For each benchmark in our benchmark suite, we ran every mode on it with a
timeout of 30 seconds, and averaged the result over 5 runs.
Figure~\ref{fig:times} summarizes the results of these tests. We find that our
algorithm is able to synthesize all of the benchmarks in our benchmark suite in
under 30 seconds. Without compositional synthesis, the synthesis algorithm is
able to solve 26 problem instances. Recall from
Section~\ref{subsec:optimizations} that users may tune the termination
metric. Success on this set of benchmarks requires we tune that termination
codition (e.g. for some benchmarks, allowing the system flexibility to spend
more time considering more of the search space). In
Section~\ref{subsec:effect-of-heuristics}, we will consider the effect of fixing
the termination metric. \saz{If there's space: Maybe say something about how hard it was to tune?}

@@

To compare to the existing bijective synthesis algorithm, we run our symmetric
synthesis algorithm on the original Optician benchmark suite. \afm{8} of these
benchmarks had to be slightly altered to work with our new system. We altered
these benchmarks by either providing additional examples or tuning termination parameters.
Without these alterations, the symmetric synthesis yielded a lens that fit
the specification but that was not the desired lens.

@@

For each benchmark in our benchmark suite, we ran every mode on it with a
timeout of 30 seconds, and averaged the result over 5 runs.
Figure~\ref{fig:times_bijective} summarizes the results of these tests. On
average, \SSOpt{} took 2.1 times (\afm{3.5} seconds) longer to complete than
\BSOpt{}. The slowest completed benchmark for both synthesis algorithms is
\texttt{xml\_to\_augeas.boom}, a benchmark that converts arbitrary XML up to
depth 3 into a serialized version of the structured dictionary representation
used in Augeas. This benchmark takes 18.9 seconds for the symmetric synthesis
algorithm to complete, and 9.3 seconds the bijective synthesis algorithm to
complete.

@@

\subsection{The Effects of Our Heuristics}
\label{subsec:effect-of-heuristics}
We evaluate the usefulness of (1) our information theoretic metric, and (2) our
termination heuristic.  To this end, we run our program in several different modes:\\
\begin{tabulary}{\linewidth}{rL}
  \AnyOpt{}: & We synthesize lenses without use of the information
               theoretic preference metric.  Once \GreedySynth finds a lens, the
  synthesis procedure returns that lens.\\
  \FLOpt{}: &  The first time \GreedySynth finds a lens, the synthesis procedure
              returns that lens. However, \GreedySynth uses the information
              theoretic preference metric to choose a lens.\\
  \CCOpt{}: & We replace our information theoretic cost metric with one where
              the cost of the lens is the number of disonnects plus the number
              of merges.\\
  \ZTOpt{}: & \ZTOpt{} runs the algorithm with the termination parameter
              universally set to 0 (default). \\
  \TFTOpt{}: & \TFTOpt{} runs the algorithm with the
               termination parameter universally set to 25. \\
  \NTFTOpt{}: & \NTFTOpt{} runs the algorithm with the
                termination parameter universally set to -25. \\
\end{tabulary}

@@

We experimented with the \CCOpt{} mode to determine whether the complexity of the information theoretic measure was really needed. Related work on string transformations has often used simpler measures such as ``avoid constants'' which align with, but are simpler than our measures. The \CCOpt{} mode is an example of such a simple measure---it operates by counting disconnects, which put a complete stop to information transfer, like constants, and merges, which eliminate the information in a union. Figure~\ref{fig:metric} summarizes the result of this experiment. We find that
our metrics are very important for finding the correct lens. Our
information-theoretic metric is critical for finding the correct lens. Indeed,
only \afm{10} of our benchmarks succeeded when running in \CCOpt{} mode.
